---
title: "Do L2 Spanish acquire scalar implicatures?"
subtitle: "DS4L final project"
author: "Robert Esposito"
institute: "Rutgers State University"
date: "2016/12/12 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
# Introduction

Replication of [Not just algunos, but indeed unos L2ers can acquire scalar implicatures in L2 Spanish](https://www.researchgate.net/profile/Roumyana-Slabakova/publication/311884036_Not_just_algunos_but_indeed_unos_L2ers_can_acquire_scalar_implicatures_in_L2_Spanish_Studies_in_honour_of_Juana_M_Liceras/links/595e9e7a0f7e9b8194b7228a/Not-just-algunos-but-indeed-unos-L2ers-can-acquire-scalar-implicatures-in-L2-Spanish-Studies-in-honour-of-Juana-M-Liceras.pdf) using simulated data

```{r include=FALSE}
source(
  here::here("scripts","libraries.R")
)
```


---
# Introduction

Do L2 Spanish speakers acquire monolingual-like use of scalar implicatures?

Context: Only three out of four dogs bark at the postman.

  a.    Unos perros ladraron al cartero.
  
  b.    Algunos perros ladraron al cartero
  
Context: All four dogs bark at the postman.

  a.    Unos perros ladraron al cartero.
  
  b.    # Algunos perros ladraron al cartero
  
---
# Research question
1.    Do L2 Spanish speakers calculate the "some, but not all" scalar implicature from *algunos*? 

2.    Do L2 Spanish speakers differ from monolingual Spanish speakers in their acceptability judgements about algunos/unos?
---
# Materials & Methods

-   200 monolingual Spanish, 200 L2 Spanish
-   Exposed to videos with four characters
    -   3/4 characters vs 4/4 characters do an action
    -   felicitous vs infelicitous statement about the video
-   Participants watch video then say whether a statement about the video is felicitous (binary response)
---
# Materials & Methods
## Conditions
```{r echo=FALSE}
tibble(
  word = c("unos","unos","unos","unos","algunos","algunos","algunos","algunos"),
  felicitous = c("felicitous","felicitous","infelicitous","infelicitous","felicitous","felicitous","infelicitous","infelicitous"),
  condition = c("3/4", "4/4","3/4","4/4","3/4","4/4","3/4","4/4"),
  n = c(5,5,5,5,5,5,5,5)
) %>% kable()
```
---
# Descriptive Results
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="500px"}
include_graphics(
  here(
    "slides","plot1.png"
  )
)
```
---
```{r message=FALSE, warning=FALSE, include=FALSE}
source(
  here(
    "scripts","model.R"
  )
)
```
# Statistical Model
-   Linear Mixed Model
    -   Criterion: % acceptance
    -   Fixed effects: group (mono, L2), number (3/4 or 4/4 characters doing an action), word (algunos, unos)
    -   Random effect: participant
---
# Statistical Model
```{r include=FALSE}
anova(mod0,mod1,mod2,mod3,mod4)
```
The following model performed best in nested model (p < 0.001)

% acceptance ~ mono + word + condition + (1 | id)
---
# Statistical Model
```{r echo=FALSE}
summary(mod3)$coefficients %>%
  kable()
```
---
# Discussion
-   L2 Spanish perform similarly to monolingual Spanish speakers
-   L2 Spanish speakers distinguish the scalar implicature of *algunos*
---